---
title: "Earnings"
author: "Luiz H. Monticelli"
date: "Apr. 15, 2024"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

# Libraries

```{r setup, include=FALSE, cache=F}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

```{r}
library(tidyverse) 
library(dplyr) 
library(ggplot2) 
library(knitr) 
library(psych)
library(tidycensus)
```

# Variables

```{r}
v2022 <- load_variables(2022, "acs1", cache = TRUE)
```

## Row names

Getting row names **for all** computer, engineering, and science occupations from v2022 to later join by name:

```{r}
names_male_oc <- v2022 %>%
  filter(name >= "B24020_019" & name <= "B24020_034")

names_male_oc[[2]] <- sub("^Estimate!!Total:!!Male:!!Management, business, science, and arts occupations:!!", "", names_male_oc[[2]])

names_male_oc <- names_male_oc  %>% 
  select(1,2) %>%
  rename(variable = name)
```

```{r}
names_female_oc <- v2022 %>%
  filter(name >= "B24020_170" & name <= "B24020_185")

names_female_oc[[2]] <- sub("^Estimate!!Total:!!Female:!!Management, business, science, and arts occupations:!!", "", names_female_oc[[2]])

names_female_oc <- names_female_oc  %>% 
  select(1,2) %>%
  rename(variable = name)
```

## Values

Creating a vector **male_oc** with all male computer, engineering, and science occupations:

```{r}
male_oc <- c("B24020_019", "B24020_020", "B24020_021", "B24020_022", "B24020_023", "B24020_024", "B24020_025", "B24020_026", "B24020_027", "B24020_028", "B24020_029", "B24020_030", "B24020_031", "B24020_032", "B24020_033", "B24020_034")
```

Creating a vector **female_oc** with all female computer, engineering, and science occupations:

```{r}
female_oc <- c("B24020_170", "B24020_171", "B24020_172", "B24020_173", "B24020_174", "B24020_175", "B24020_176", "B24020_177", "B24020_178", "B24020_179", "B24020_180", "B24020_181", "B24020_182", "B24020_183", "B24020_184", "B24020_185")
```

## Getting data from 2010 to 2022 (skipping 2020)

```{r}
male_oc_data <- list()

for (year in 2010:2022) {
  if (year == 2020) { #Skipping crappy 2020
    next
  }
  
  male_oc_data[[as.character(year)]] <- get_acs(
    geography = "state",
    survey = "acs1",
    variables = male_oc,
    year = year
  )
}

combined_male_oc_data <- bind_rows(male_oc_data, .id = "Year") %>%
  mutate(sex = "Male") %>%
  inner_join(names_male_oc, by="variable")
```

```{r}
female_oc_data <- list()

for (year in 2010:2022) {
  if (year == 2020) { #Skipping crappy 2020
    next
  }
  
  female_oc_data[[as.character(year)]] <- get_acs(
    geography = "state",
    survey = "acs1",
    variables = female_oc,
    year = year
  )
}

combined_female_oc_data <- bind_rows(female_oc_data, .id = "Year") %>%
  mutate(sex = "Female") %>%
  inner_join(names_female_oc, by="variable")
```

### One table to rule them all

```{r}
occupation_by_state <- rbind(combined_male_oc_data, combined_female_oc_data)
```

## Prettying

```{r}
occupation_by_state <- occupation_by_state %>% 
  select(1,3,4,7,8,5) %>%
  rename(State = NAME, Sex = sex, Occupation = label, Estimate = estimate, Variable = variable)
```

```{r}
## write.csv(occupation_by_state, "occupation.csv", row.names = FALSE)
```

# Master table = occupation_by_state

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX

## Getting down to business

Are there any correlations with election results?

```{r}
filtered_data <- occupation_by_state %>%
  filter(grepl("B24020_019|B24020_170", Variable, ignore.case = TRUE))
```

```{r}
## write.csv(filtered_data, "filtered.csv", row.names = FALSE)
```

```{r}
male_data <- filtered_data %>% filter(Sex == "Male")
female_data <- filtered_data %>% filter(Sex == "Female")

# Reshape each dataframe
male_data <- male_data %>%
  pivot_wider(names_from = Sex, values_from = Estimate) %>%
  rename(male_value = Male)

female_data <- female_data %>%
  pivot_wider(names_from = Sex, values_from = Estimate) %>%
  rename(female_value = Female)

# Merge the reshaped dataframes
combined_data <- left_join(male_data, female_data, by = c("Year", "State"))
```

```{r}
combined_data <- combined_data %>% 
  select(1,2,4,5,8)
```

```{r}
combined_data <- combined_data %>%
  mutate(
    total_value = male_value + female_value,
    percent_male_value = round((male_value / total_value) * 100, 2),
    percent_female_value = round((female_value / total_value) * 100, 2)
  )
```

```{r}
combined_data <- combined_data %>% 
  select(1,2,3,7,8) %>%
  rename(Occupation = Occupation.x, Pct_Male = percent_male_value, Pct_Female = percent_female_value)
```

Descriptives

```{r}
describe(combined_data)
```

-\> Mean and Median are almost the same.

# 2022 - Current landscape

The states with the highest employment gap between male and female population are:

```{r}
an_2022 <- combined_data %>% 
  filter(Year == 2022) %>% 
  filter(State != "Puerto Rico")
```

Does it correlate with the last Presidential Election result?

```{r}
library(readr)
elections2020 <- read_csv("elections2020.csv")
```

```{r}
an_2022 <- left_join(an_2022, elections2020, by = "State")
```

### Histograms

```{r}
hist(an_2022$Pct_Male)
hist(an_2022$Pct_Female)
hist(an_2022$Biden)
hist(an_2022$Trump)
```

### Plot

```{r}
plot(an_2022$Trump, an_2022$Pct_Male)
```

### Correlation

```{r}
cor <- round(cor(an_2022[4:7], use="pairwise.complete.obs"),2)
View(cor)
```

## Regression

```{r}
options(scipen = 999)
mod1 <- lm(Trump ~ Pct_Male, data=an_2022)
summary(mod1)
```

## -\> We found a strong correlation (0.68) between percentage of votes to Trump and the occupation gap between male and female in STEM fields. The Adjusted R-squared (0.449) means that 44.9% of the inequality can be explained by political stance. We know the R-squared is statistically significant because of the p-value that's well below 0.05. For every percent in support to the Republican candidate we expect we expect the likelihood of inequality to go up by 2.55%.
